---
title: "Chapter 4"
author: "Martine"
date: "24/9/2019"
output: html_document
---


```{r}
library(tidyverse)

facebookData <- read.delim("FacebookNarcissism.dat", header = TRUE)

```

If we don’t like the circles then we can change the shape of the points by specifying this for
the geom. shape = 17 

```{r}
ggplot(facebookData, aes(x = NPQC_R_Total, y = Rating)) + 
  geom_point(shape = 6)
```
Atthis stage we don’t know whether a rating represented coolness, attractiveness or whatever.
It would be nice if we could differentiate different ratings, perhaps by plotting them in
different colours. We can do this by setting the colour aesthetic to be the variable Rating_Type.

```{r}
ggplot(facebookData, aes(x = NPQC_R_Total, y = Rating)) + 
  geom_point(aes(colour=Rating_Type)) 

```


```{r}
ggplot(facebookData, aes(x = NPQC_R_Total, y = Rating)) +
  geom_point(aes(colour = Rating_Type), position = "jitter")
```

```{r}
ggplot(facebookData, aes(x = NPQC_R_Total, y = Rating)) +
  geom_point(aes(shape = Rating_Type), position = "jitter")
```

```{r}

```

